On change of priors in one-vs-one classification methods

### 2025-05-13 theory.tex
Fix Assumption 1 and notation of binary classifiers

\begin{assumption} \label{ass:1}
For all $x$ in the feature space we have $d_i(x) > 0$.
\end{assumption}
Commonly used classifiers such as linear discriminant analysis, (penalized) logistic regression or support vector machines all satisfy this assumption.

@@

\begin{align}
	p_{ij}^\Pi(x) = \frac{d_i(x) P_i}{d_i(x)P_i + d_j(x)P_j}.
\end{align}

Knowledge of vector $\boldsymbol{p}^\Pi(x)$ is equivalent to knowing the functions $p^\Pi_{ij}(x)$. Indeed under Assumption \ref{ass:1} we have
\begin{align}
p_{ij}^\Pi(x) = \frac{p_i^\Pi(x)}{p_i^\Pi(x) + p_j^\Pi(x)}. \label{eq:bt1}
\end{align}

@@

\begin{prop}[Fundamental theorem of one-vs-one classification]
	 \label{prop:binary2multi}
	 Knowledge of binary Bayes classifiers $p_{ij}^\Pi$ determines the multiclass Bayes classifier $\boldsymbol{p}^\Pi$.
\end{prop}
\begin{proof}
	Given a multiclass distribution $\boldsymbol{p} = (p_1, \ldots,p_K)$ on $K$ classes, let $\alpha_{ij}(\boldsymbol{p})$ denote the odds of $C_i$ to $C_j$:
	\begin{align}
		\alpha_{ij}(\boldsymbol{p}) = \frac{p_i}{p_j} \label{eq:alpha}
	\end{align}

	From \eqref{eq:bt1} we can deduce the odds of classes $C_j$ to $C_i$ for the Bayes classifier since
	\begin{align}
	\alpha_{ij}(\boldsymbol{p}^\Pi) = \frac{p_j^\Pi(x)}{p_j^\Pi(x)} = \frac{1}{p^\Pi_{ij}(x)}- 1.
	\end{align}
	Knowing the odds of $C_2$ to $C_1$, $C_3$ to $C_2$ up to $C_{K}$ to $C_{K-1}$ together with the requirement that probabilities sum to one uniquely determines the whole vector $\boldsymbol{p}^\Pi(x)$.
\end{proof}
%\begin{proof}
%Indeed, we have 
%\begin{align*}
%\frac{1}{r_{ij}(x)} - 1 = \frac{p_j^\Pi}{p_i^\Pi}
%\end{align*}
%\end{proof}

@@

Equations forming system \eqref{eq:bt2} are called Bradley-Terry equations.
The system is usually inconsistent and additional assumptions are needed to construct an estimator. A method to solve Bradley-Terry equations is called a \emph{coupling method}. Its input can be represented by a matrix $\boldsymbol{R}(x)$ with off-diagonal entries being $r_{ij}(x)$. Schematically, classification using coupling method $\boldsymbol{v}$ proceeds as indicated in Figure \ref{fig:coupling}.

@@

>>>>

\begin{assumption} \label{ass:1}
For all $x$ in the feature space we have $d_i(x) > 0$.
\end{assumption}

@@

\begin{align}
	p_{ij}(x) = \frac{d_i(x) P_i}{d_i(x)P_i + d_j(x)P_j}.
\end{align}

Knowledge of vector $\boldsymbol{p}^\Pi(x)$ is equivalent to knowing the functions $p^\Pi_{ij}(x)$. Indeed under Assumption \ref{ass:1} we have
\begin{align}
p_{ij}(x) = \frac{p_i^\Pi(x)}{p_i^\Pi(x) + p_j^\Pi(x)}. \label{eq:bt1}
\end{align}

@@

\begin{prop}[Fundamental theorem of one-vs-one classification]
	 \label{prop:binary2multi}
	 Knowledge of binary Bayes classifiers $p_{ij}$ determines the multiclass Bayes classifier $\boldsymbol{p}^\Pi$.
\end{prop}
\begin{proof}
	Given a multiclass distribution $\boldsymbol{p} = (p_1, \ldots,p_K)$ on $K$ classes, let $\alpha_{ij}(\boldsymbol{p})$ denote the odds of $C_i$ to $C_j$:
	\begin{align}
		\alpha_{ij}(\boldsymbol{p}) = \frac{p_i}{p_j} \label{eq:alpha}
	\end{align}

	From \eqref{eq:bt1} we can deduce the odds of classes $C_j$ to $C_i$ for the Bayes classifier since
	\begin{align}
	\alpha_{ij}(\boldsymbol{p}^\Pi) = \frac{p_j^\Pi(x)}{p_j^\Pi(x)} = \frac{1}{p_{ij}(x)}- 1.
	\end{align}
	Knowing the odds of $C_2$ to $C_1$, $C_3$ to $C_2$ up to $C_{K}$ to $C_{K-1}$ together with the requirement that probabilities sum to one uniquely determines the whole vector $\boldsymbol{p}^\Pi(x)$.
\end{proof}
%\begin{proof}
%Indeed, we have 
%\begin{align*}
%\frac{1}{r_{ij}(x)} - 1 = \frac{p_j^\Pi}{p_i^\Pi}
%\end{align*}
%\end{proof}

@@

Equations forming system \eqref{eq:bt2} are called Bradley-Terry equations. In parallel to Assumption \ref{ass:1} it is customary to make the following assumption on estimates $r_{ij}$.

\begin{assumption} \label{ass:2}
For any $x$ in the feature space $r_{ij}(x) > 0$.
\end{assumption}

Commonly used classifiers such as linear discriminant analysis, (penalized) logistic regression or support vector machines all satisfy this assumption.

A method to solve Bradley-Terry equations is called a \emph{coupling method}. Its input can be represented by a matrix $\boldsymbol{R}(x)$ with off-diagonal entries being $r_{ij}(x)$. Schematically, classification using coupling method $\boldsymbol{v}$ proceeds as indicated in Figure \ref{fig:coupling}.

@@

The system of Bradley-Terry equations is usually inconsistent and additional assumptions are needed to construct an estimator. In Sections \ref{sec:des:exact} and \ref{sec:des:inexact} we review various approaches to construct coupling methods.
